Introducing Statistical Design of Experiments to SPARQL Endpoint Evaluation

### 2013-05-08 exp1/iswc2013.tex
small changes in the large fractional
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@529 43ef1190-90fd-4dd7-a769-d0121f02b80e

The importance of finding flaws in the levels is highlighted by the
normal plot in figure~\ref{fig:frac64hugenormal}. The smaller
experiment made it feasible to run with a larger data-set, this
experiment uses the original 15~MTriples data-set of
\cite{mxro:Morsey2011DBpedia}. The picture is now as hinted in the
previous discussion, for larger ``TripleC'', our simulated performance
is actually worse, as evident from the positive
``Implement''. We also see that the interactions ``Implement:Lang''
and ``Implement:TripleC'' is highly significant. It is clear what
happened: For larger ``TripleC'', the delay we introduced in the
langMatches function has a devastating effect, since it delays the
execution for every returned triple, and thereby completely dominates
the execution time.

Moreover, we note that the ``Machine'' factor has become highly
significant in the other end along with its ``Machine:TripleC''
interaction. One should take careful notice that the ``Machine''
factor encompasses many different things, that should be different
factors; CPU, disks, available RAM, installed software, tuning
parameters and so on. It is advantageous to have such broad factors in
the design in early experiments, because they may point out unexpected
effects that should be further scrutinized. The ``Machine'' factor is
a good example of usage that also helps us manage
complexity: If possible, one may group many factors into one if one is
in doubt of its relevance and break it up in subsequent
experiments. In this case, the cause of the performance problem was
highly audible: It was the old P-ATA disks of the larger machine, but
in general, such clues are not available to the experimenter.

@@

Note that we do not take advantage of the usual performance of 4store
in this simulation, to the contrary, we chose 4store because of its
relative simplicity to modify it to suite our needs for illustrative
purposes, which consists of degrading it in many ways. The small
data-sets used here does not imply that the approach will not work for
large data-sets; properly configured, actual optimization will give
very different runtimes, and the small number of runs will make it
possible to evaluate a very large number of factors.

>>>>

The importance of finding flaws in the levels is highlighted by the
normal plot in figure~\ref{fig:frac64hugenormal}. The smaller
experiment made it feasible to run with a larger data-set, this
experiment uses the original 15~MTriples data-set of
\cite{mxro:Morsey2011DBpedia}. The picture is now as hinted in the
previous discussion, for larger ``TripleC'', our simulated performance
is actually worse, as evident from the positive ``Implement''. We also
see that the interactions ``Implement:Lang'' and ``Implement:TripleC''
is highly significant. It is clear what happened: For larger
``TripleC'', the delay we introduced in the langMatches function has a
devastating effect, since it delays the execution for every triple
that the query engine applies the filter to, and thereby completely
dominates the execution time.

Moreover, we note that the ``Machine'' factor has become highly
significant in the other end along with its ``Machine:TripleC''
interaction. One should take careful notice that the ``Machine''
factor encompasses many different things, that should be different
factors; CPU, disks, available RAM, installed software, tuning
parameters and so on. It is advantageous to have such broad factors in
the design in early experiments, because they may point out unexpected
effects that should be further scrutinized. The ``Machine'' factor is
a good example of usage that also helps us manage complexity: If
possible, one may group many factors into one if one is in doubt of
its relevance and break it up in subsequent experiments if deeper
understanding is needed. In this case, the cause of the performance
problem was highly audible: It was the old P-ATA disks of the larger
machine, but in general, such clues are not available to the
experimenter, therefore more factors are needed.

@@

Note that we do not take advantage of the usual performance of 4store
in this simulation, to the contrary, we chose 4store because of its
relative simplicity to modify it to suite our needs for illustrative
purposes, which consists of degrading it in several ways. The small
data-sets used here does not imply that the approach will not work for
large data-sets; properly configured, actual optimization will give
very different runtimes, and the small number of runs will make it
possible to evaluate a very large number of factors.
